Why is it that statistical tests for residuals are not widely used? An explanation using visual inference.

### 2023-01-20 paper_comparison/paper_comparison/paper_comparison.Rmd
changed % discussion for the non-linear example

For lineups produced by the non-linearity model, conventional tests reject `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% count(conv_reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% 
and visual tests reject `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% of the time. Of the tests rejected by the conventional test, `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% filter(conv_reject == "Reject") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`%, so approximately half as many as the conventional test. Surprisingly, the visual test rejects `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% filter(conv_reject == "Not") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% of lineups where the conventional test does not reject.

are rejected by both the conventional and visual tests, while 42% are accepted by both tests. Thus, 69% of the time two tests will agree with each other. There are 26% of lineups rejected by conventional tests only.

Interestingly, in spite of the greater power of conventional tests, 7% of the lineups are rejected by visual tests only. Recall that the RESET test used in our analysis includes powers of fitted values up to four. However, the "M" and the "Triple-U" shapes constructed from the Hermite polynomials contain power of the regressor up to six and 18 respectively. In fact, by including the power of fitted values up to six in the RESET test, all these cases will be rejected. We also can observe from the example plot displayed in Figure \ref{fig:poly-example} that panel three is distinctly different from others, exhibiting a "M" shape. This clearly suggests users of the RESET test and any other residual-based conventional tests that require the specification of variables of interest to check the residual plot before conducting the test, such that the correct choice of variables can be made.

>>>>

For lineups produced by the non-linearity model, conventional tests reject `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% count(conv_reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% 
and visual tests reject `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% of the time. Of the tests rejected by the conventional test, `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% filter(conv_reject == "Reject") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% are rejected by the visual test, that is, approximately half as many as the conventional test. Surprisingly, the visual test rejects `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% filter(conv_reject == "Not") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% (i.e. 10 out of 70) of lineups where the conventional test does not reject.

Figure \ref{fig:poly-example} shows one of the ten lineups where the visual test rejected the residual plot. The data plot, in position three is very obvious, and different from the nulls. The visual test was correct to reject. To understand why this is, one needs to return to the way the RESET test is applied. It requires a parameter indicating degree of polynomial to test for, and the recommendation is to generically use 4 (XXX REF). However, the "M" and the "Triple-U" shapes constructed from the Hermite polynomials use power up to 6 and 18 respectively. All 10 lineups rejected by the visual test but not the RESET were of types "M" or "Triple-U". If the RESET test had been applied using the power up to 6, all the decisions for these lineups would have been to reject. The recommendation of the polynomial power for the RESET should be revised, perhaps. This illustrates the sensitivity of conventional testing to the parameters, and it also points to a limitation that one needs to know the data structure in order to set the parameters for the test.

<!--We also can observe from the example plot displayed in Figure \ref{fig:poly-example} that panel three is distinctly different from others, exhibiting a "M" shape. This clearly suggests users of the RESET test and any other residual-based conventional tests that require the specification of variables of interest to check the residual plot before conducting the test, such that the correct choice of variables can be made.-->

### 2023-01-20 paper_comparison/paper_comparison/paper_comparison.tex
changed % discussion for the non-linear example

For lineups produced by the non-linearity model, conventional tests
reject 51\% and visual tests reject 32\% of the time. Of the tests
rejected by the conventional test, 49\%, so approximately half as many
as the conventional test. Surprisingly, the visual test rejects 14\% of
lineups where the conventional test does not reject.

are rejected by both the conventional and visual tests, while 42\% are
accepted by both tests. Thus, 69\% of the time two tests will agree with
each other. There are 26\% of lineups rejected by conventional tests
only.

Interestingly, in spite of the greater power of conventional tests, 7\%
of the lineups are rejected by visual tests only. Recall that the RESET
test used in our analysis includes powers of fitted values up to four.
However, the ``M'' and the ``Triple-U'' shapes constructed from the
Hermite polynomials contain power of the regressor up to six and 18
respectively. In fact, by including the power of fitted values up to six
in the RESET test, all these cases will be rejected. We also can observe
from the example plot displayed in Figure \ref{fig:poly-example} that
panel three is distinctly different from others, exhibiting a ``M''
shape. This clearly suggests users of the RESET test and any other
residual-based conventional tests that require the specification of
variables of interest to check the residual plot before conducting the
test, such that the correct choice of variables can be made.

>>>>

For lineups produced by the non-linearity model, conventional tests
reject 51\% and visual tests reject 32\% of the time. Of the tests
rejected by the conventional test, 49\% are rejected by the visual test,
that is, approximately half as many as the conventional test.
Surprisingly, the visual test rejects 14\% (i.e.~10 out of 70) of
lineups where the conventional test does not reject.

Figure \ref{fig:poly-example} shows one of the ten lineups where the
visual test rejected the residual plot. The data plot, in position three
is very obvious, and different from the nulls. The visual test was
correct to reject. To understand why this is, one needs to return to the
way the RESET test is applied. It requires a parameter indicating degree
of polynomial to test for, and the recommendation is to generically use
4 (XXX REF). However, the ``M'' and the ``Triple-U'' shapes constructed
from the Hermite polynomials use power up to 6 and 18 respectively. All
10 lineups rejected by the visual test but not the RESET were of types
``M'' or ``Triple-U''. If the RESET test had been applied using the
power up to 6, all the decisions for these lineups would have been to
reject. The recommendation of the polynomial power for the RESET should
be revised, perhaps. This illustrates the sensitivity of conventional
testing to the parameters, and it also points to a limitation that one
needs to know the data structure in order to set the parameters for the
test.
